---
title: "Exercise 9"
author: "Claire Brandes"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: leonids
    toc: yes
---

# Exercise 9: Practice Simple Linear Regression

# Step 1

Loading data

```{r}
library(tidyverse)

f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"
d <- read_csv(f, col_names = TRUE)
```

Exploratory data analysis (median, min, max, 1st and 3rd quartiles, mean, sd)

```{r}
library(skimr)
d <- na.omit(d)
my_skim <- skim_with(numeric = sfl( 
                                   p0 = NULL,
                                   p50 = NULL,
                                   p100 = NULL,
                                   hist = NULL,
                                   median, 
                                   min, 
                                   max, 
                                   p25 = ~ quantile(., probs = .25), 
                                   p75 = ~quantile(., probs = .75),
                                   mean, 
                                   sd)) 
my_skim(d, where(is.numeric))
```

## Step 2

Plot brain size as a function of social group size, longevity, juvenile period length, and reproductive lifespan

```{r}
sgs_ECV <- ggplot(d, aes(x = Group_size, y = ECV)) + 
  geom_point() + 
  theme_classic() +
  xlab("Social Group Size") + ylab("Brain Size")

Lon_ECV <- ggplot(d, aes(x = Longevity, y = ECV)) +
  geom_point() +
  theme_classic() +
  xlab("Longevity") + ylab("Brain Size")

jpl_ECV <- ggplot(d, aes(x = Weaning, y = ECV)) +
  geom_point() +
  theme_classic() +
  xlab("Juvenile Period Length") + ylab("Brain Size")

repro_ECV <- ggplot(d, aes(x = Repro_lifespan, y = ECV)) +
  geom_point() +
  theme_classic() +
  xlab("Reproductive Lifespan") + ylab("Brain Size")

library(ggpubr)
ggarrange(sgs_ECV, Lon_ECV, jpl_ECV, repro_ECV)
```

## Step 3

Derive by hand the ordinary least squares regression coefficients $\beta1$ and $\beta0$ for ECV as a function of social group size

```{r}
s3 <- d %>% filter(!is.na(Group_size), !is.na(ECV))
(b1 <- cov(s3$Group_size, s3$ECV)/var(s3$Group_size))
(b0 <- mean(s3$ECV) - (b1*mean(s3$Group_size)))
```

## Step 4

Confirm that you get the same results using the lm() function

```{r}
(m <- lm(ECV ~ Group_size, data = s3))
```

## Step 5

Repeat the analysis above for catarrhines, platyrrhines, and strepsirrhines separately. 

for Catarrhines:
```{r}
cat <- s3 %>%
  filter(Taxonomic_group == "Catarrhini")

lm(ECV ~ Group_size, data = cat)
```

for Platyrrhines:

```{r}
plat <- s3 %>%
  filter(Taxonomic_group == "Platyrrhini")

lm(ECV ~ Group_size, data = plat)
```

for Strepsirhines:

```{r}
strep <- s3 %>%
  filter(Taxonomic_group == "Strepsirhini")

lm(ECV ~ Group_size, data = strep)
```

The coefficients looks like they differ among the three groups. To determine if they are meaningfully different I could generate a test statistic for observed differences in coefficient values, generate a null distribution of coefficient values, and permute group assignments for each line of data. 

## Step 6

For the first regression of ECV on social group size, calculate the standard error for the slope coefficient, the 95% CI, and the *p* value associated with this coefficient by hand.

```{r}
# Calculating error sum of squares
SSE <- sum((m$model$ECV - m$fitted.values)^2)

# Calculating error mean sum of squares
MSE <- SSE/74

# Calculating sum squares of group size 
SSX <- sum((m$model$Group_size - mean(m$model$Group_size))^2)

# Calculating standard error in B1 (slope coefficient)
(se_b1 <- sqrt(MSE/SSX))
```

Calculating p-value and CI: step incomplete 

Test statistic =  group_size estimate / group_size std. error 

```{r}
2.874 / 0.5732771
```

Extract same information from the results of running the lm() function:

```{r}
library(broom)
tidy(m)
```

